
---
title: "Traffic Death Exploratory Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    collapse: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
    slidy_presentation:
    duration: 60
    footer: '2020'
#  pdf_document:
#    number_sections: true
#    toc: yes
#    toc_depth: '3'
---
<!-- 
  R -e 'rmarkdown::render(input="cut.Rmd",output_format = "slidy_presenation")'  
-->
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(eval=TRUE, collapse=FALSE, echo = TRUE, warning = FALSE, message = FALSE, tidy=TRUE,tidy.opts=list(width.cutoff=50))
library(gganimate)
library(tidyverse)
data <- read.csv('~/Downloads/nassCDS.csv')
library(tidyverse)
```

#Create dataframe to find deaths by vehicle year
```{r}
averageAcc <- data %>% 
  select(yearVeh, yearacc, injSeverity, dead) %>% 
  group_by(yearVeh) %>% drop_na() %>%
  summarise(avgSev = mean(injSeverity), deadCount = sum(dead == "dead"),aliveCount = sum(dead == "alive"), deadPCT = deadCount /(deadCount + aliveCount) *100) %>% filter(aliveCount + deadCount > 50)

```

#Create dataframe to find deaths by safety device usage
```{r}
safetyDeaths <- data %>% 
  select(airbag, seatbelt, injSeverity, dead) %>% 
  group_by(airbag,seatbelt) %>% drop_na() %>% filter(injSeverity <= 5) %>%
  summarise(avgSev = mean(injSeverity), deadCount = sum(dead == "dead"),aliveCount = sum(dead == "alive"), deadPCT = deadCount /(deadCount + aliveCount) *100) %>% filter(aliveCount + deadCount > 50)

```

#Create dataframe to find deaths by DVCat rating
```{r}
DVCATData <- data %>% 
  select(dvcat,frontal, injSeverity, dead) %>% 
  group_by(dvcat,frontal) %>% drop_na() %>% filter(injSeverity <= 5) %>%
  summarise(avgSev = mean(injSeverity), deadCount = sum(dead == "dead"),aliveCount = sum(dead == "alive"), deadPCT = deadCount /(deadCount + aliveCount) *100) %>% filter(aliveCount + deadCount > 50)

DVCATData2 <- data %>% 
  select(dvcat,frontal, injSeverity, dead, airbag, seatbelt) %>% 
  group_by(dvcat,frontal,seatbelt,airbag) %>% drop_na() %>% filter(injSeverity <= 5) %>%
  summarise(avgSev = mean(injSeverity), deadCount = sum(dead == "dead"),aliveCount = sum(dead == "alive"), deadPCT2 = deadCount /(deadCount + aliveCount) *100) %>% filter(aliveCount + deadCount > 50)
```

#Create dataframe to find deaths by demographic
```{r}
averageDemo <- data %>% 
  select(yearacc, injSeverity, dead,weight,sex,occRole,ageOFocc) %>% 
  group_by(yearacc,occRole) %>% drop_na() %>%
  summarise(avgSev = mean(injSeverity), deadCount = sum(dead == "dead"),aliveCount = sum(dead == "alive"), deadPCT = deadCount /(deadCount + aliveCount) *100)

```

#Dataframe for age/weight
```{r}
averageDemo <- data %>% 
  select(yearacc, injSeverity, dead,weight,sex,occRole,ageOFocc) %>% 
  group_by(yearacc,occRole) %>% drop_na() %>%
  summarise(avgSev = mean(injSeverity), deadCount = sum(dead == "dead"),aliveCount = sum(dead == "alive"), deadPCT = deadCount /(deadCount + aliveCount) *100)
```

#Graphs for severity/death based on vehicle year
```{r}
ggplot(averageAcc, aes(x = yearVeh, y=deadPCT)) +
  theme_dark(base_size = 10) + geom_point(size=2, shape=15) 

ggplot(averageAcc, aes(x = yearVeh, y=avgSev)) +
  theme_dark(base_size = 10) + geom_point(size=2, shape=15) 
```

#Graphs for severity/death based on safety usage
```{r}
ggplot(safetyDeaths, aes(x = airbag, y = deadPCT)) +
     geom_bar(stat = "identity") + facet_grid(~seatbelt)

ggplot(safetyDeaths, aes(x = airbag, y = avgSev)) +
     geom_bar(stat = "identity") + facet_grid(~seatbelt)
```

#Graphs for severity/death based on DVCat ratings
```{r}
ggplot(DVCATData, aes(x = dvcat, y = deadPCT)) +
     geom_bar(stat = "identity") + facet_grid(~frontal)

ggplot(DVCATData2, aes(x = dvcat, y = deadPCT2)) +
     geom_bar(stat = "identity") + facet_grid(~seatbelt)
```

#Graph for severity/death based on occupant role 
```{r}
ggplot(averageDemo, aes(x = occRole , y=deadPCT)) +
         geom_bar(stat = "identity")
```
  
  
  

  